Understanding Factors Contributing to a Successful Himalayan Expedition

### 2024-04-29 paper/paper.qmd

The Himalayan Database is a compilation of records for all expeditions that have climbed in the Nepal Himalaya. The database is based on the expedition archives of Elizabeth Hawley, a longtime journalist based in Kathmandu, and it is supplemented by information gathered from books, alpine journals and correspondence with Himalayan climbers.

@@

The three datasets I considered included information about all peaks in Nepal, all expeditions on those peaks and all members on those expeditions. I then combines the data from these three datasets to form the main analysis dataset.

@@

Winter expeditions have lower success probabilities, probably owning to the extreme cold conditions at higher altitudes. The interaction of these two factors provide an interesting insight. It is observed that the success probabilities during different seasons for peaks between 5400 metres and 5749 metres are all comparatively closer to each other than compared to peaks between 7500 metres and 7849 metres in height. All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in winter to around 0.6 in autumn or spring.

@@

While age-related decline in mountaineering performance is noticeable, it is essential to recognize that success in mountaineering is not solely determined by physical capabilities. Experience, skill, preparation, and mindset also play an important roles in summit success. Older climbers can compensate for physiological limitations through greater experience, strategic planning, and mental resilience, enabling them to pursue their mountaineering goals despite age-related challenges.

@@

Additionally, while I focused on focused on a few factors, there are numerous other variables that could impact the likelihood of a successful summit. Nationality and group size were two factors which were available in the raw data and I believed were factors which could have played a significant difference. These were later removed due to the lack of consistency in those variable’s. For instance, the way the number of members were reported differed from record to record. As the dataset also included the people summiting the other side of border peaks, some countries reported this number differently than others. For expeditions in Nepal, this was usually the number of foreigners listed on the expedition permit whereas in China, this was usually the number of non-hired members.

@@

>>>>

The Himalayan Database is a compilation of records for all expeditions that have climbed in the Nepal Himalaya. The database is based on the expedition archives of Elizabeth Hawley, a long-time journalist based in Kathmandu, and it is supplemented by information gathered from books, alpine journals and correspondence with Himalayan climbers.

@@

The three datasets I considered included information about all peaks in Nepal, all expeditions on those peaks and all members on those expeditions. I then combine the data from these three datasets to form the main analysis dataset.

@@

Winter expeditions have lower success probabilities, probably owning to the extreme cold conditions at higher altitudes. The interaction of these two factors provide an interesting insight. It is observed that the success probabilities during different seasons for peaks between 5400 metres and 5749 metres are all comparatively closer to each other than compared to peaks between 7500 metres and 7849 metres in height. All seasons' success probabilities for peaks between 5400 metres and 5749 metres fall between 0.85 to 0.9 whereas for peaks between 7500 metres and 7848 metres, the success probabilities go from around 0.35 in winter to around 0.6 in autumn or spring.

@@

While age-related decline in mountaineering performance is noticeable, it is essential to recognize that success in mountaineering is not solely determined by physical capabilities. Experience, skill, preparation, and mindset also play important roles in summit success. Older climbers can compensate for physiological limitations through greater experience, strategic planning, and mental resilience, enabling them to pursue their mountaineering goals despite age-related challenges.

@@

Additionally, while I focused on focused on a few factors, there are numerous other variables that could impact the likelihood of a successful summit. Nationality and group size were two factors which were available in the raw data and I believed were factors which could have played a significant difference. These were later removed due to the lack of consistency in those variables. For instance, the way the number of members were reported differed from record to record. As the dataset also included the people summiting the other side of border peaks, some countries reported this number differently than others. For expeditions in Nepal, this was usually the number of foreigners listed on the expedition permit whereas in China, this was usually the number of non-hired members.

@@

\appendix
